Equalizer 2.0 --\\Convergence of a Parallel Rendering Framework

### 2018-01-18 paperIEEE.tex
git-svn-id: https://dev.ifi.uzh.ch/vmml/svn/Equalizer16@91 db7c848b-7755-4df2-b763-8796797afa52

To benchmark the data distribution we used two datasets: The david statue at 2mm
resolution (\fig{fDavid2mm}) and 3D volumes of a spike frequencies of an
electrical simulation of three million neurons (\fig{fVolumes}).

@@

RLE compression has a very low overhead, but does not compress the data, since
it only removes ``blank space'' in the data. The Snappy compression, used as
default in \textsf{Collage}, achieves the same compression ratio as the LZ
variants at a much higher speed. The ZStandard compressor has roughly the same
speed as the LZ variants at the lowest compression level, but provides
significantly better compression. At higher compression levels it can improve
the compression ratio slightly, but at a high cost for the compression speed.

@@

The ply data is little compressible, the default compressor achieves a 10\%
reduction. This is due to a high entropy in the data and the dominant use of
floating point values. Overall the profile is similar to the generic benchmark,
at a smaller compression rate.

The volume data on the other hand is sparsely populated and using integer (byte
and short) values, which is easier to compress. The naive RLE implementation
already achieves a good compression, showing that the smaller volume contains at
most 28\% empty space and the bigger volume at most 43\%. Snappy and ZStandard
can reduce the spike data much further, reducing the data to a few megabytes.
Surprisingly, the beechnut data set does not yield significant higher
compression with the modern Snappy and ZStandard libraries.

@@

In this section we present how data distribution and synchronization performs in
real-world applications. We extracted the existing data distribution code in a
PLY (eqPly) and a volume renderer (Livre) into a benchmark application to
measure the time to initially map all the objects on the render client nodes,
and to perform a commit-sync of the full data set after mapping has been
established. All figures observe a noticable measurement jitter due to other
services running on the cluster. The details of the benchmark algorithm can be
found in the implementation~\cite{eqObjectBench}. We used the same data sets as
in the previous section. After eight processes, a nodes start to run two
processes per node, which share CPU, memory and network interface bandwidth.

Object mapping is measured using the following settings: \textbf{none}
distributes the raw, uncompressed, and unbuffered data, \textbf{compression}
uses the Snappy compressor to distribute unbuffered data, \textbf{buffered}
reuses uncompressed, serialized data for mappings of other nodes, and
\textbf{compression buffered} compresses this buffer. Unbuffered operations need
to reserialize, and potentially recompress, the master object data for each
slave node. Each slave instance needs to deserialize and decompress the data.

@@

The David statue at 2mm resolution is organized in a kD-tree for rendering. Each
node is a separate distributed object, which has two child node objects. A
total of $1023$ objects is distributed and synchronized. \fig{fDavidDist} shows
the distribution times for this data set. Due to limited compressibility of the
data, the results are relative similar. Compressing the data repeatedly for
each client leads to decreased performance, since the compression overhead
cannot be amortized by the decreased transmission time. Buffering data slightly
improves performance by reducing the CPU and copy overhead. Combining
compression and buffering leads to the best performance, although only by about
10\%.

During synchronization, data is pushed from the master process to all mapped
slaves using a unicast connection to each slave. While the results in the
\fig{fDavidDist} (middle) are relatively close to each other, we can still
observe how the tradeoff between compression ratio and speed influences overall
performance. Better, slower compression algorithms lead to better overall better
performance when amortized over many send operations.

The volume data sets are distributed in a single object, serializing the
raw volume buffer. The spike volume data set has a significant compression
ratio, which reflects in the results in \fig{fSpikesDist}. Compression for this
data is benefitial for transmitting data over a 10~Gb/s link even for a single
slave process only (970 MB/s vs 570 MB/s raw data transfer). Buffering has
little benefit since the serialization of volume data is trivial. Buffered
compression makes a huge difference, since the compression cost can be amortized
over $n$ nodes, reaching raw data transmission rates of 3.7 GB/s with the
default Snappy compressor and at best 4.4 GB/s with ZStandard at level 1.

The distribution of the beechnut data set also behaves as expected
(\fig{fBeechnutDist}). Due to the larger object size, uncompressed transmission
is slightly faster compared to the spike data set at 700 MB/s, and compressed
transmission does not improve the mapping performance, likely due to increased
memory pressure caused by the large data size. The comparison of the various
compression engines is consistent with the benchmarks in
\fig{fCompressorDetail}; RLE, Snappy and the LZ variants are very close to each
other, and ZSTD1 can provide better performance after four nodes due to the
better compression ratio.

Finally, we compare data distribution speed using different protocols. In this
benchmark, data synchronization time of the spike volume data set is measured,
as in \fig{fSpikesDist} (middle). Buffering is enabled, and compression is
disabled to focus on the raw network performance. \fig{fIFDist} shows the
performance using various protocols.

@@

\begin{figure*}[p!]\center
  \includegraphics[height=.19\textheight]{images/daviddist}
  \caption{\label{fDavidDist}Object mapping (left) and data synchronization
    time (middle, detail view right) for the David 2mm data set in
    \fig{fDavid2mm}}
\end{figure*}
\begin{figure*}[p!]\center
  \includegraphics[height=.19\textheight]{images/spikesdist}
  \caption{\label{fSpikesDist}Object mapping (left) and data synchronization
    time (middle, detail view right) for the spike data set in
    \fig{fVolumes}, left}
\end{figure*}
\begin{figure*}[p!]\center
  \includegraphics[height=.19\textheight]{images/beechnutdist}
  \caption{\label{fBeechnutDist}Object mapping (left) and data synchronization
    time (middle, detail view right) for the beechnut data set in
    \fig{fVolumes}, right}
\end{figure*}
\begin{figure*}[p!]\center
  \includegraphics[height=.19\textheight]{images/ifdist}
  \caption{\label{fIFDist}Object synchronization using different network
    transports}
\end{figure*}

>>>>

To benchmark the data distribution we used two datasets: The david statue at 2mm
resolution (\fig{fDavid2mm}) and 3D volumes of spike frequencies of an
electrical simulation of three million neurons (\fig{fVolumes}).

@@

RLE compression has a very low overhead but
% does not compress the data, since it only
merely removes ``blank space'' in the data. The Snappy compression, used as
default in \textsf{Collage}, achieves the same compression ratio as the LZ
variants at a much higher speed. The ZStandard compressor has roughly the same
speed as the LZ variants at the lowest compression level, but provides
significantly better compression. At higher compression levels it can improve
the compression ratio slightly, but at a high cost for the compression speed.

@@

The ply data is little compressible, the default compressor achieves a 10\%
reduction. This is due to a high entropy in the data and the dominant use of
floating point values. Overall, the profile is similar to the generic benchmark,
at a smaller compression rate.

The volume data on the other hand is sparsely populated and using integer (byte
and short) values, which is easier to compress. The naive RLE implementation
already achieves a good compression rate, showing that the smaller volume contains at
most 28\% empty space and the bigger volume at most 43\%. Snappy and ZStandard
can reduce the spike data much further, reducing the data to a few megabytes.
Surprisingly, the beechnut data set does not yield significantly higher
compression with the modern Snappy and ZStandard libraries.

@@

In this section we present how data distribution and synchronization performs in
real-world applications. We extracted the existing data distribution code in a
PLY (eqPly) and a volume renderer (Livre) into a benchmark application to
measure the time to initially map all the objects on the render client nodes,
and to perform a commit-sync of the full data set after mapping has been
established. All figures observe a noticable measurement jitter due to other
services running on the cluster. The details of the benchmark algorithm can be
found in the implementation~\cite{eqObjectBench}. We used the same data sets as
in the previous section. After eight processes, a node starts to run two
processes per node, which share CPU, memory and network interface bandwidth \FIXME{unclear}.

Object mapping is measured using the following settings: \textbf{none}
distributes the raw, uncompressed, and unbuffered data, \textbf{compression}
uses the Snappy compressor to distribute unbuffered data, \textbf{buffered}
reuses uncompressed, serialized data for mappings of other nodes, and
\textbf{compression buffered} compresses the buffer holding the data \FIXME{unclear}. Unbuffered operations need
to reserialize, and potentially recompress, the master object data for each
slave node. Each slave instance needs to deserialize and decompress the data.

@@

The David statue at 2mm resolution is organized in a kD-tree for rendering. Each
node is a separate distributed object, which has two child node objects. A
total of $1023$ objects is distributed and synchronized. \fig{fDavidDist} shows
the distribution times for this data set. Due to limited compressibility of the
data, the results are relatively similar. Compressing the data repeatedly for
each client leads to decreased performance, since the compression overhead
cannot be amortized by the decreased transmission time. Buffering data slightly
improves performance by reducing the CPU and copy overhead. Combining
compression and buffering leads to the best performance, although only by about
10\%.

During synchronization, data is pushed from the master process to all mapped
slaves using a unicast connection to each slave. While the results in the
\fig{fDavidDist} (middle) are relatively close to each other, we can still
observe how the tradeoff between compression ratio and speed influences overall
performance. Better, slower compression algorithms lead to improved overall
performance when amortized over many send operations.

The volume data sets are distributed in a single object, serializing the
raw volume buffer. The Spike volume data set has a significant compression
ratio, which is reflected by the results in \fig{fSpikesDist}. Compression for this
data is beneficial for transmitting data over a 10~Gb/s link even for a single
slave process only (970 MB/s vs 570 MB/s raw data transfer) \FIXME{unclear}. Buffering has
little benefit since the serialization of volume data is trivial. Buffered
compression makes a significant difference, since the compression cost can be amortized
over $n$ nodes, reaching raw data transmission rates of 3.7 GB/s with the
default Snappy compressor and at best 4.4 GB/s with ZStandard at level 1.

The distribution of the beechnut data set also behaves as expected
(\fig{fBeechnutDist}). Due to the larger object size, uncompressed transmission
is slightly faster compared to the Spike data set at 700 MB/s, and compressed
transmission does not improve the mapping performance, likely due to increased
memory pressure caused by the large data size. The comparison of the various
compression engines is consistent with the benchmarks in
\fig{fCompressorDetail}; RLE, Snappy and the LZ variants are very close to each
other, and ZSTD1 can provide better performance after four nodes due to the
better compression ratio.

Finally, we compare data distribution speed using different protocols. In this
benchmark, data synchronization time of the Spike volume data set is measured,
as in \fig{fSpikesDist} (middle). Buffering is enabled, and compression is
disabled to focus on the raw network performance. \fig{fIFDist} shows the
performance using various protocols.

@@

\begin{figure*}[p!]\center
  \includegraphics[height=.19\textheight]{images/daviddist}
  \caption{\label{fDavidDist}Object mapping (left) and data synchronization
    time (middle, detail view right) for the David 2mm data set in
    \fig{fDavid2mm}}
\end{figure*}
\begin{figure*}[p!]\center
  \includegraphics[height=.19\textheight]{images/spikesdist}
  \caption{\label{fSpikesDist}Object mapping (left) and data synchronization
    time (middle, detail view right) for the Spike data set in
    \fig{fVolumes}, left}
\end{figure*}
\begin{figure*}[p!]\center
  \includegraphics[height=.19\textheight]{images/beechnutdist}
  \caption{\label{fBeechnutDist}Object mapping (left) and data synchronization
    time (middle, detail view right) for the beechnut data set in
    \fig{fVolumes}, right}
\end{figure*}
\begin{figure*}[p!]\center
  \includegraphics[height=.19\textheight]{images/ifdist}
  \caption{\label{fIFDist}Object synchronization using different network
    transports}
\end{figure*}
